---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit THAT file, then run devtools::build_readme(). -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)
library(roperators)
```

# roperators

<!-- badges: start -->
[![CRAN version](https://img.shields.io/cran/v/roperators?style=flat-square)](https://cran.r-project.org/package=roperators)
[![metacran downloads](https://cranlogs.r-pkg.org/badges/last-month/roperators)](https://cran.r-project.org/package=roperators)
[![License: MIT](https://img.shields.io/cran/l/roperators?style=flat-square)](https://cran.r-project.org/package=roperators)
[![R-CMD-check](https://github.com/BenWiseman/roperators/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/BenWiseman/roperators/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

> Small, friendly operators and helpers that let your R code read the way you already think it.

R is a wonderful language to work in — but here and there it asks you to write
something a little clunkier than the thought in your head. `roperators` quietly
smooths those moments over. There's nothing magic going on underneath: just a
handful of small, well-behaved operators and helpers, all in base R, that make
everyday code calmer to read and write — and a gentle welcome for folks arriving
from Python and elsewhere.

(It's pronounced *rop-er-ators*, by the way — not *r-operators*.)

## Installation

```r
# the released version, from CRAN:
install.packages("roperators")

# or the development version, from GitHub:
# install.packages("remotes")
remotes::install_github("BenWiseman/roperators")
```

```r
library(roperators)
```

## String arithmetic

Let's start with the one nearly everyone misses coming from other languages —
adding strings together:

```{r}
"using infix (%) operators " %+% "lets R do string addition"

"abcabc" %-% "c"            # subtract a pattern
"ha" %s*% 3                 # multiply
"an apple a day" %s/% "a"   # divide: how many times does the pattern appear?
```

## In-place modifiers (*à la* `+=`)

We've all written the same long line twice just to nudge a column up by one. You
needn't:

```{r}
x <- 1
x %+=% 2
x

d <- iris
d$Sepal.Length[d$Species == "setosa"] %+=% 1   # add 1, in place
```

There's a whole little family — `%-=%`, `%*=%`, `%/=%`, `%^=%`, `%root=%`,
`%log=%` — plus `%regex=%` (apply a regex in place), `%regex<-%` (overwrite
matching elements), and `%na<-%` (gently fill in the `NA`s):

```{r}
x <- c(1, NA, 3, NA)
x %na<-% 0
x
```

## Comparisons that behave

`NA` and floating point trip up almost everyone at some point — and honestly,
it's not your fault, it's just the defaults. Here's a softer landing:

```{r}
c(1, NA, 3) == c(1, NA, 4)    # base R: the NA leaks through
c(1, NA, 3) %==% c(1, NA, 4)  # roperators: NA == NA is treated as TRUE

(0.1 + 0.1 + 0.1) == 0.3      # base R: FALSE (floating point)
(0.1 + 0.1 + 0.1) %~=% 0.3    # roperators: TRUE

5 %><% c(1, 10)   # strictly between
1 %>=<% c(1, 10)  # between, inclusive
```

You also get `%===%` (strict value-and-class equality, like JavaScript's `===`),
`%>=%` / `%<=%` (with `NA` equality), and `%>~%` / `%<~%` (greater/less than *or*
approximately equal).

## Logical & SQL-style pattern matching

```{r}
"z" %ni% c("a", "b", "c")          # not in
TRUE %xor% FALSE                   # exclusive or
c("FOO", "bar") %rlike% "foo"      # case-insensitive LIKE
```

## ✨ New in 1.4

A few new friends arrived in this release:

```{r}
# f-strings: interpolate expressions straight into a string
name <- "Ben"; n <- 2
f("Hi {name}, you have {n} message{if (n != 1) 's'}")

# %else%: a calm fallback if an expression decides to error
sqrt("not a number") %else% NA_real_

# %/0%: safe division, so a stray zero never poisons your means
10 %/0% 0

# %+-%: a tolerance interval that drops right into the 'between' operators
4.9 %><% (5 %+-% 0.5)

# %~%: forgiving, case/whitespace-insensitive string equality
"  Yes " %~% "yes"

# as.percent(): proportions, dressed up as tidy percentages
as.percent(c(0.1, 0.005, 2/3))
```

## Shorter conversions & gentle checks

Shorter names for the conversions you reach for all the time:

```{r}
int("42"); chr(42); num("4.2"); bool("TRUE")   # as.integer(), as.character(), ...

# no more as.numeric(as.character(factor_var)):
f.as.numeric(factor(c(11, 22, 33)))

# one tidy question instead of five:
is.bad_for_calcs(c(1, NA, Inf, NaN))
```

There's a matching set of `is.*_or_null()` predicates too, along with
`is.scalar()`, `is.constant()`, `is.binary()`, and friends — handy for keeping
your guard clauses short.

## A little drawer of everyday helpers

```{r}
get_1st_word("Ada Lovelace")           # and get_last_word / get_nth_word
get_most_frequent(c("a", "b", "b"))    # the mode(s)
paste_oxford("Tom", "Dick", "Harry")   # Oxford-comma joining, done for you

mean_cc(c(1, 2, NA))                   # *_cc = the base function with na.rm = TRUE
```

…plus OS checks (`get_os()`, `is.os_mac()`, …), file-extension checks
(`is_csv_file()`, …), and `read.tsv()` / `read.psv()`.

## Learn more

* `vignette("nicer_roperators")` — the unhurried, guided tour.
* Full reference: <https://benwiseman.github.io/roperators/>

A gentle heads-up on names: a few are shared on purpose with the wider world
(`%+%` with ggplot2, `%like%`-style matching with data.table). If you've got
those loaded too, just reach for the namespaced form (`roperators::%+%`) where it
matters, and everyone gets along fine.
