---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit THAT file, then run devtools::build_readme(). -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)
library(roperators)
```

# roperators

<!-- badges: start -->
[![CRAN version](https://img.shields.io/cran/v/roperators?style=flat-square)](https://cran.r-project.org/package=roperators)
[![metacran downloads](https://cranlogs.r-pkg.org/badges/last-month/roperators)](https://cran.r-project.org/package=roperators)
[![License: MIT](https://img.shields.io/cran/l/roperators?style=flat-square)](https://cran.r-project.org/package=roperators)
[![R-CMD-check](https://github.com/BenWiseman/roperators/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/BenWiseman/roperators/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

> Small, friendly operators and helpers that let your R code read the way you already think it.

R is a wonderful language to work in ŌĆö but here and there it asks you to write
something clunkier than the thought in your head. `roperators` smooths those
moments over. Nothing magic underneath: just a handful of small, well-behaved
operators and helpers, all in base R ŌĆö and a friendly welcome for folks arriving
from Python and elsewhere.

```{r}
"foo" %+% "bar"               # add strings
(0.1 + 0.1 + 0.1) %~=% 0.3    # floating-point equality that works
c(1, NA) %==% c(1, NA)        # NA == NA treated as TRUE
```

(It's pronounced *rop-er-ators*, not *r-operators*.)

## Installation

```r
# released version, from CRAN:
install.packages("roperators")

# or the development version, from GitHub:
# install.packages("remotes")
remotes::install_github("BenWiseman/roperators")
```

```{r}
library(roperators)

"hello " %+% "world"   # your first win
```

## Find the right tool

<div id="rop-finder">
  <p><em>On the <a href="https://benwiseman.github.io/roperators/">package
  website</a> this is an interactive finder ŌĆö search by what you're trying to do
  (ŌĆ£join two stringsŌĆØ, ŌĆ£missing valuesŌĆØ, ŌĆ£read a fileŌĆØŌĆ”) and jump straight to the
  docs. Either way, the <a href="https://benwiseman.github.io/roperators/reference/index.html">full
  reference</a> lists everything.</em></p>
</div>

*New here? Take the [5-minute guided tour](https://benwiseman.github.io/roperators/articles/nicer_roperators.html), or skim the tour of features below.*

## String arithmetic

Let's start with the one nearly everyone misses coming from other languages ŌĆö
adding strings together:

```{r}
"using infix (%) operators " %+% "lets R do string addition"

"abcabc" %-% "c"            # subtract a pattern
"ha" %s*% 3                 # multiply
"an apple a day" %s/% "a"   # divide: how many times does the pattern appear?
```

## In-place modifiers (*├Ā la* `+=`)

We've all written the same long line twice just to nudge a column up by one. You
needn't:

```{r}
x <- 1
x %+=% 2
x

d <- iris
d$Sepal.Length[d$Species == "setosa"] %+=% 1   # add 1, in place
```

There's a whole family ŌĆö `%-=%`, `%*=%`, `%/=%`, `%^=%`, `%root=%`, `%log=%` ŌĆö
plus `%regex=%` (apply a regex in place), `%regex<-%` (overwrite matching
elements), and `%na<-%` (fill in the `NA`s):

```{r}
x <- c(1, NA, 3, NA)
x %na<-% 0
x
```

## Comparisons that behave

`NA` and floating point trip up almost everyone ŌĆö it's not your fault, just the
defaults:

```{r}
c(1, NA, 3) == c(1, NA, 4)    # base R: the NA leaks through
c(1, NA, 3) %==% c(1, NA, 4)  # roperators: NA == NA is treated as TRUE

(0.1 + 0.1 + 0.1) == 0.3      # base R: FALSE (floating point)
(0.1 + 0.1 + 0.1) %~=% 0.3    # roperators: TRUE

5 %><% c(1, 10)   # strictly between
1 %>=<% c(1, 10)  # between, inclusive
```

You also get `%===%` (strict value-and-class equality, like JavaScript's `===`),
`%>=%` / `%<=%` (with `NA` equality), and `%>~%` / `%<~%` (greater/less than *or*
approximately equal).

## Logical & SQL-style pattern matching

A few logical conveniences, plus SQL-style pattern matching:

```{r}
"z" %ni% c("a", "b", "c")          # not in
TRUE %xor% FALSE                   # exclusive or
c("FOO", "bar") %rlike% "foo"      # case-insensitive LIKE
```

## f-strings & safe everyday operators

A few of these arrived in 1.4 (see the [changelog](https://benwiseman.github.io/roperators/news/index.html)):

```{r}
# f-strings: interpolate expressions straight into a string
name <- "Ben"; n <- 2
f("Hi {name}, you have {n} message{if (n != 1) 's'}")

# %else%: a fallback when an expression errors
sqrt("not a number") %else% NA_real_

# %/0%: safe division, so a stray zero never poisons your means
10 %/0% 0

# %+-%: a tolerance interval that drops right into the 'between' operators
4.9 %><% (5 %+-% 0.5)

# %~%: forgiving, case/whitespace-insensitive string equality
"  Yes " %~% "yes"

# as.percent(): proportions, dressed up as tidy percentages
as.percent(c(0.1, 0.005, 2/3))
```

## Shorter conversions & quick checks

Shorter names for the conversions you reach for all the time:

```{r}
int("42"); chr(42); num("4.2"); bool("TRUE")   # as.integer(), as.character(), ...

# no more as.numeric(as.character(factor_var)):
f.as.numeric(factor(c(11, 22, 33)))

# TRUE where a value is NA, NaN, or Inf:
is.bad_for_calcs(c(1, NA, Inf, NaN))
```

There's a matching set of `is.*_or_null()` predicates too, along with
`is.scalar()`, `is.constant()`, `is.binary()`, and friends ŌĆö handy for keeping
your guard clauses short.

## Everyday helpers

```{r}
get_1st_word("Ada Lovelace")           # and get_last_word / get_nth_word
get_most_frequent(c("a", "b", "b"))    # the mode(s)
paste_oxford("Tom", "Dick", "Harry")   # Oxford-comma joining

mean_cc(c(1, 2, NA))                   # *_cc = the base function with na.rm = TRUE
```

ŌĆ”plus OS checks (`get_os()`, `is.os_mac()`, ŌĆ”), file-extension checks
(`is_csv_file()`, ŌĆ”), and `read.tsv()` / `read.psv()`.

## At a glance

| You wantŌĆ”                       | Reach for                           |
|---------------------------------|-------------------------------------|
| String concat / subtract        | `%+%` / `%-%`                       |
| String repeat / count           | `%s*%` / `%s/%`                     |
| In-place maths                  | `%+=%` `%-=%` `%*=%` `%/=%` `%^=%`  |
| Fill NAs / regex edit in place  | `%na<-%` / `%regex=%` / `%regex<-%` |
| NA-aware (in)equality           | `%==%` `%>=%` `%<=%`                |
| Floating-point equality         | `%~=%` `%>~%` `%<~%`                |
| Strict (value + class) equality | `%===%`                            |
| Between (excl / incl)           | `%><%` / `%>=<%`                   |
| Not-in / xor / all-or-nothing   | `%ni%` / `%xor%` / `%aon%`         |
| String interpolation            | `f()`                              |
| Inline error fallback           | `%else%`                           |
| Safe divide / tolerance         | `%/0%` / `%+-%`                    |
| Fuzzy string match              | `%~%`                              |

## Next steps

* **[Take the 5-minute guided tour ŌåÆ](https://benwiseman.github.io/roperators/articles/nicer_roperators.html)**
* **[Browse all functions ŌåÆ](https://benwiseman.github.io/roperators/reference/index.html)**

A quick note on shared names: `%+%` is also ggplot2's, and `%like%`-style
matching is data.table's. If you load those too, use the namespaced form
(`roperators::%+%`) where it matters.
