---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit THAT file, then run devtools::build_readme(). -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)
library(roperators)
```

# roperators

<!-- badges: start -->
[![CRAN version](https://img.shields.io/cran/v/roperators?style=flat-square)](https://cran.r-project.org/package=roperators)
[![metacran downloads](https://cranlogs.r-pkg.org/badges/last-month/roperators)](https://cran.r-project.org/package=roperators)
[![License: MIT](https://img.shields.io/cran/l/roperators?style=flat-square)](https://cran.r-project.org/package=roperators)
[![R-CMD-check](https://github.com/BenWiseman/roperators/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/BenWiseman/roperators/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

> Additional operators and helpers to make R code easier to **read, write, and maintain** — and a friendlier landing for people arriving from Python and other languages.

`roperators` gives you the small things you keep wishing R had: string
arithmetic (`'foo' + 'bar'`), in-place modifiers (`x += 1`), comparisons that
don't fall over on `NA` or floating point, and a pile of tiny quality-of-life
helpers. Pure base R, no heavy dependencies.

(It's pronounced *rop-er-ators*, not *r-operators*.)

## Installation

```r
# Released version from CRAN:
install.packages("roperators")

# Development version from GitHub:
# install.packages("remotes")
remotes::install_github("BenWiseman/roperators")
```

```r
library(roperators)
```

## String arithmetic

```{r}
"using infix (%) operators " %+% "lets R do string addition"

"abcabc" %-% "c"            # subtract a pattern
"ha" %s*% 3                 # multiply
"an apple a day" %s/% "a"   # divide: how many times does the pattern occur?
```

## In-place modifiers (*à la* `+=`)

No more `df$x[long$condition] <- df$x[long$condition] + 1`:

```{r}
x <- 1
x %+=% 2
x

d <- iris
d$Sepal.Length[d$Species == "setosa"] %+=% 1   # add 1, in place
```

You also get `%-=%`, `%*=%`, `%/=%`, `%^=%`, `%root=%`, `%log=%`,
`%regex=%` (apply a regex in place), `%regex<-%` (overwrite matching elements),
and `%na<-%` (fill in the `NA`s):

```{r}
x <- c(1, NA, 3, NA)
x %na<-% 0
x
```

## Comparisons that behave

`NA`-aware equality, real floating-point equality, and `between`:

```{r}
c(1, NA, 3) == c(1, NA, 4)    # base R: NA leaks through
c(1, NA, 3) %==% c(1, NA, 4)  # roperators: NA == NA is TRUE

(0.1 + 0.1 + 0.1) == 0.3      # base R: FALSE (floating point)
(0.1 + 0.1 + 0.1) %~=% 0.3    # roperators: TRUE

5 %><% c(1, 10)   # strictly between
1 %>=<% c(1, 10)  # between, inclusive
```

Plus `%===%` (strict value-and-class equality, like JavaScript's `===`),
`%>=%` / `%<=%` (with `NA` equality), and `%>~%` / `%<~%` (greater/less than
*or* approximately equal).

## Logical & SQL-style pattern matching

```{r}
"z" %ni% c("a", "b", "c")          # not in
TRUE %xor% FALSE                   # exclusive or
c("FOO", "bar") %rlike% "foo"      # case-insensitive LIKE
```

## ✨ New in 1.4

```{r}
# f-strings: interpolate expressions straight into a string
name <- "Ben"; n <- 2
f("Hi {name}, you have {n} message{if (n != 1) 's'}")

# %else%: inline fallback if an expression errors
sqrt("not a number") %else% NA_real_

# %/0%: safe division (no Inf/NaN sneaking into your means)
10 %/0% 0

# %+-%: tolerance interval that composes with the 'between' operators
4.9 %><% (5 %+-% 0.5)

# %~%: fuzzy, case/whitespace-insensitive string equality
"  Yes " %~% "yes"

# as.percent(): proportions -> tidy percentage strings
as.percent(c(0.1, 0.005, 2/3))
```

## Shorter conversions & safety checks

```{r}
int("42"); chr(42); num("4.2"); bool("TRUE")   # as.integer(), as.character(), ...

# stop writing as.numeric(as.character(factor_var)):
f.as.numeric(factor(c(11, 22, 33)))

# one guard instead of five:
is.bad_for_calcs(c(1, NA, Inf, NaN))
```

There are matching `is.*_or_null()` predicates, `is.scalar()`, `is.constant()`,
`is.binary()`, and friends.

## Everyday helpers

```{r}
get_1st_word("Ada Lovelace")           # and get_last_word / get_nth_word
get_most_frequent(c("a", "b", "b"))    # mode(s)
paste_oxford("Tom", "Dick", "Harry")   # Oxford-comma joining

mean_cc(c(1, 2, NA))                   # *_cc = the base function with na.rm = TRUE
```

…plus OS checks (`get_os()`, `is.os_mac()`, …), file-extension checks
(`is_csv_file()`, …), and `read.tsv()` / `read.psv()`.

## Learn more

* `vignette("nicer_roperators")` — the guided tour.
* Full reference: <https://benwiseman.github.io/roperators/>

A quick note on collisions: a few names are deliberately shared with the wider
ecosystem (`%+%` with ggplot2, `%like%`-style matching with data.table). If you
load those packages too, reach for the namespaced form (e.g.
`roperators::%+%`) where it matters.
