---
title: "roperators: cleaner, friendlier R"
output:
  prettydoc::html_pretty:
    toc: true
    theme: cayman
    highlight: github
vignette: >
  %\VignetteIndexEntry{roperators: cleaner, friendlier R}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo     = TRUE,
  collapse = TRUE,
  comment  = "#>"
)
library(roperators)
```

`roperators` adds the small things you keep wishing base R had: string
arithmetic, in-place modifiers, comparisons that survive contact with `NA` and
floating point, and a pile of tiny quality-of-life helpers. It's pure base R
with no heavy dependencies, and it's designed to be especially welcoming to
people arriving from Python and other languages.

This vignette is a guided tour. If you just want the highlights, here they are:

```{r}
"foo" %+% "bar"               # string addition
(0.1 + 0.1 + 0.1) %~=% 0.3    # floating-point equality that works
c(1, NA) %==% c(1, NA)        # NA == NA is TRUE here

name <- "you"
f("hello {name}, 2 + 2 = {2 + 2}")   # f-strings!
```

# String arithmetic

One of the most common gripes from people coming to R is the lack of string
arithmetic. So we added it.

```{r}
my_string <- "using infix (%) operators " %+% "lets R do string addition"
my_string

# subtraction removes a pattern
my_string %-% "lets R do string addition"

# multiplication repeats (%*% was already taken, so it's %s*%)
"ha" %s*% 3
```

And something you *can't* do in Python — **string division**, which counts how
many times a pattern appears (regular expressions welcome):

```{r}
"an apple a day keeps the malignant spirit of Steve Jobs at bay" %s/% "a"

# with a regular expression
"an apple a day keeps the malignant spirit of Steve Jobs at bay" %s/% "Steve Jobs|apple"
```

# In-place modifiers (*à la* `+=`)

How many times have you written something like
`df$x[long$condition] <- df$x[long$condition] + 1`? The line doesn't even fit on
the page. With `roperators`:

```{r}
x <- 1
x %+=% 2
x

d <- iris
# add 1 to setosa sepal lengths, in place
d$Sepal.Length[d$Species == "setosa"] %+=% 1
```

The full set is `%+=%`, `%-=%`, `%*=%`, `%/=%`, `%^=%`, `%root=%`, and `%log=%`.
`%+=%` and `%-=%` also work on strings:

```{r}
x <- "ab"
x %+=% "c"
x
```

## Replacing missing values and regex matches

`%na<-%` fills in the `NA`s, and `%regex=%` / `%regex<-%` edit in place:

```{r}
x <- c(NA, 1, 2, 3)
x %na<-% 0
x

x <- c("a1b", "b1", "c", "d0")
x %regex=% c("\\d+", "#")   # replace the matched part
x
```

# Comparisons that behave

## `NA`-aware equality

An `NA` doesn't technically equal another `NA` — but most of the time, for
your purposes, it should. How many `if` statements have broken on this?

```{r}
a <- c(NA, "foo", "foo", NA)
b <- c(NA, "foo", "bar", "bar")

a == b      # base R: NA leaks through
a %==% b    # roperators: NA == NA is TRUE
```

`%>=%` and `%<=%` carry the same NA-aware behaviour.

## Floating-point equality

The classic trap. Innocent statistics students are seldom warned:

```{r}
(0.1 + 0.1 + 0.1) == 0.3    # FALSE (!)
(0.1 + 0.1 + 0.1) %~=% 0.3  # TRUE

# greater/less-than-or-approximately-equal
(0.1 + 0.1 + 0.1) %>~% 0.3
(0.1 + 0.1 + 0.1) %<~% 0.3
```

## Between, and strict equality

```{r}
5 %><% c(1, 10)    # strictly between
1 %>=<% c(1, 10)   # inclusive
5 %><% c(10, 1)    # reversed bounds are fine too

# %===% is strict value-AND-class equality, like JavaScript's ===
x <- int(2)
x == 2       # TRUE
x %===% 2    # FALSE (different class)
x %===% int(2)
```

# Logical and SQL-style operators

```{r}
"z" %ni% c("a", "b", "c")    # not in
TRUE %xor% FALSE             # exclusive or
TRUE %aon% TRUE              # all-or-nothing: both TRUE or both FALSE

# SQL-style LIKE
c("FOO", "bar", "fizz") %rlike% "foo"   # case-insensitive
c("dOe", "doe")         %perl% "[a-z]O" # case-sensitive, Perl regex
```

# ✨ New in 1.4

A small, opinionated set of additions.

**`f()` — string interpolation (R's f-strings).** Anything in `{ }` is evaluated
in the calling environment:

```{r}
who <- "Ben"; n <- 2
f("Hi {who}, you have {n} new message{if (n != 1) 's'}")
f("today's first letters: {head(LETTERS, n)}")   # vectors are collapsed
```

**`%else%` — inline fallback** when an expression might error (the fallback is
only evaluated if needed):

```{r}
sqrt("not a number") %else% NA_real_
(1:3)[[99]]          %else% "out of range"
```

**`%/0%` — safe division** that returns `NA` instead of letting `Inf`/`NaN`
poison a downstream `sum()` or `mean()`:

```{r}
c(10, 20, 30) %/0% c(2, 0, 5)
```

**`%+-%` — a tolerance interval** that drops straight into the `between`
operators:

```{r}
5 %+-% 0.5
4.9 %><% (5 %+-% 0.5)
```

**`%~%` — fuzzy string equality** (ignores case and whitespace), the string
sibling of `%~=%`:

```{r}
"  Yes " %~% "yes"
c("Apple", "PEAR") %~% c("apple", "pear")
```

**`as.percent()` — proportions as tidy strings:**

```{r}
as.percent(c(0.1, 0.005, 2 / 3))
as.percent(2 / 3, digits = 0)
```

# Shorter type conversions

R's conversion syntax is wordy. These help:

```{r}
chr(42)      # as.character()
int(42.9)    # as.integer()
num("4.2")   # as.numeric()
bool("TRUE") # as.logical()

# the infamous factor-to-number gotcha, solved:
fac <- factor(c(11, 22, 33))
as.numeric(fac)     # 1 2 3  -- almost never what you want
f.as.numeric(fac)   # 11 22 33

# convert to a class chosen at run time
as.class(255, "roman")
```

# Safety and type checks

Rather than chaining five checks, ask one question:

```{r}
# would these break a calculation?
is.bad_for_calcs(c(1, NA, Inf, NaN, 5))

is.scalar(1)
is.constant(c(1, 1, 1))
is.binary(c("a", "b", "a"))
```

There's a full family of `is.*_or_null()` predicates too, handy for validating
optional function arguments.

# Everyday helpers

```{r}
# pull pieces out of vectors and strings
get_1st_word("Ada Lovelace")
get_last_word("Ada Lovelace")
get_most_frequent(c("a", "b", "b", "c", "b"))

# Oxford-comma joining
paste_oxford("Tom", "Dick", "Harry")

# complete-cases stats: just add _cc for na.rm = TRUE
mean_cc(c(1, 2, NA))
sd_cc(c(1, 2, 3, NA))

# environment checks
get_os()
get_R_version()

# file-extension checks
is_csv_file(c("a.csv", "b.txt"))
```

# Cheat sheet

| You want…                          | Use                                  |
|------------------------------------|--------------------------------------|
| String concat / subtract           | `%+%` / `%-%`                        |
| String repeat / count              | `%s*%` / `%s/%`                      |
| In-place maths                     | `%+=%` `%-=%` `%*=%` `%/=%` `%^=%`   |
| Fill NAs / regex edit in place     | `%na<-%` / `%regex=%` / `%regex<-%` |
| NA-aware (in)equality              | `%==%` `%>=%` `%<=%`                 |
| Floating-point equality            | `%~=%` `%>~%` `%<~%`                 |
| Strict (value + class) equality    | `%===%`                             |
| Between (excl / incl)              | `%><%` / `%>=<%`                    |
| Not-in / xor / all-or-nothing      | `%ni%` / `%xor%` / `%aon%`          |
| SQL-style LIKE                     | `%rlike%` / `%perl%`                |
| String interpolation               | `f()`                               |
| Inline error fallback              | `%else%`                            |
| Safe divide / tolerance            | `%/0%` / `%+-%`                     |
| Fuzzy string match                 | `%~%`                               |

# A note on collisions

A few names are deliberately shared with the wider ecosystem — `%+%` with
ggplot2, and `%like%`-style matching with data.table. If you load those packages
as well, reach for the namespaced form (`roperators::%+%`) where it matters.
